### 2019-11-27 ic-journal/multivariate.tex
Fixed some merge conflicts

Proposition~\ref{prop:transFV} \hl{is essential in the proofs of the
  multivariable versions of all unique-solution
  theorems.
In comparison with} the above proof, the proof of the analogous
result for bound variables is  easier:
\begin{proposition}
\label{prop:transBV}
if $E \arr{\mu} E'$ then $\bv{E'} \subseteq \bv{E}$, or formally:
\begin{alltt}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{E} \HOLBoundVar{u} \ensuremath{\HOLBoundVar{E}\sp{\prime}}. \HOLBoundVar{E} \HOLTokenTransBegin\HOLBoundVar{u}\HOLTokenTransEnd \ensuremath{\HOLBoundVar{E}\sp{\prime}} \HOLSymConst{\HOLTokenImp{}} \HOLConst{BV} \ensuremath{\HOLBoundVar{E}\sp{\prime}} \HOLSymConst{\HOLTokenSubset{}} \HOLConst{BV} \HOLBoundVar{E}\hfill{[TRANS_BV]}
\end{alltt}
\end{proposition}
\begin{proof}
Following the proof of Prop.~\ref{prop:transFV}, we do action
induction on $E \arr{\mu} E'$ and simplify all subgoals by set-theoretic
facts. The only subgoal left is:
\begin{alltt}
        \HOLinline{\HOLConst{BV} \ensuremath{\HOLFreeVar{E}\sp{\prime}} \HOLSymConst{\HOLTokenSubset{}} \HOLFreeVar{X} \HOLConst{INSERT} \HOLConst{BV} \HOLFreeVar{E}}
   ------------------------------------
    0.  \HOLinline{\HOLConst{BV} \ensuremath{\HOLFreeVar{E}\sp{\prime}} \HOLSymConst{\HOLTokenSubset{}} \HOLConst{BV} \ensuremath{(}\ensuremath{[}\HOLConst{rec} \HOLFreeVar{X} \HOLFreeVar{E}\ensuremath{/}\HOLFreeVar{X}\ensuremath{]} \HOLFreeVar{E}\ensuremath{)}}
\end{alltt}
Note the difference with the proof of Prop.~\ref{prop:transFV}: now
we have \texttt{INSERT} in place of \texttt{DELETE}. Then, we directly
use the assumption and the transitivity of $\subseteq$ and try to
prove the following goal instead:
\begin{alltt}
        \HOLinline{\HOLConst{BV} \ensuremath{(}\ensuremath{[}\HOLConst{rec} \HOLFreeVar{X} \HOLFreeVar{E}\ensuremath{/}\HOLFreeVar{X}\ensuremath{]} \HOLFreeVar{E}\ensuremath{)} \HOLSymConst{\HOLTokenSubset{}} \HOLFreeVar{X} \HOLConst{INSERT} \HOLConst{BV} \HOLFreeVar{E}}
   ------------------------------------
    0.  \HOLinline{\HOLConst{BV} \ensuremath{\HOLFreeVar{E}\sp{\prime}} \HOLSymConst{\HOLTokenSubset{}} \HOLConst{BV} \ensuremath{(}\ensuremath{[}\HOLConst{rec} \HOLFreeVar{X} \HOLFreeVar{E}\ensuremath{/}\HOLFreeVar{X}\ensuremath{]} \HOLFreeVar{E}\ensuremath{)}}
\end{alltt}
This is an easy goal, due to the following lemma which is
 proved by induction on $E$:
\begin{alltt}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{X} \HOLBoundVar{E} \ensuremath{\HOLBoundVar{E}\sp{\prime}}. \HOLConst{BV} \ensuremath{(}\ensuremath{[}\ensuremath{\HOLBoundVar{E}\sp{\prime}}\ensuremath{/}\HOLBoundVar{X}\ensuremath{]} \HOLBoundVar{E}\ensuremath{)} \HOLSymConst{\HOLTokenSubset{}} \HOLConst{BV} \HOLBoundVar{E} \HOLSymConst{\HOLTokenUnion{}} \HOLConst{BV} \ensuremath{\HOLBoundVar{E}\sp{\prime}}\hfill{[BV_SUBSET]}
\end{alltt}
Taking $E ' = \recu X E$ in the above lemma, we have
    \HOLinline{\HOLConst{BV}\\\;\ensuremath{(}\ensuremath{[}\HOLConst{rec}\\\;\HOLFreeVar{X}\\\;\HOLFreeVar{E}\ensuremath{/}\HOLFreeVar{X}\ensuremath{]}\\\;\HOLFreeVar{E}\ensuremath{)}\\\;\HOLSymConst{\HOLTokenSubset{}}\\\;\HOLConst{BV}\\\;\HOLFreeVar{E}\\\;\HOLSymConst{\HOLTokenUnion{}}\\\;\HOLConst{BV}\\\;\ensuremath{(}\HOLConst{rec}\\\;\HOLFreeVar{X}\\\;\HOLFreeVar{E}\ensuremath{)}},
    the right side equals to \HOLinline{\HOLConst{BV}\\\;\HOLFreeVar{E}\\\;\HOLSymConst{\HOLTokenUnion{}}\\\;\ensuremath{(}\HOLFreeVar{X}\\\;\HOLConst{INSERT}\\\;\HOLConst{BV}\\\;\HOLFreeVar{E}\ensuremath{)}} $=$
    \HOLinline{\HOLFreeVar{X}\\\;\HOLConst{INSERT}\\\;\HOLConst{BV}\\\;\HOLFreeVar{E}}. This
completes
the    proof.
\end{proof}

@@

   \HOLConst{CCS_SUBST} \HOLFreeVar{fm} \ensuremath{(}\HOLConst{rec} \HOLFreeVar{X} \HOLFreeVar{E}\ensuremath{)} \HOLTokenDefEquality{}
     \HOLKeyword{if} \HOLFreeVar{X} \HOLSymConst{\HOLTokenIn{}} \HOLConst{FDOM} \HOLFreeVar{fm} \HOLKeyword{then} \HOLConst{rec} \HOLFreeVar{X} \ensuremath{(}\HOLConst{CCS_SUBST} \ensuremath{(}\HOLFreeVar{fm} \HOLSymConst{\ensuremath{\setminus}} \HOLFreeVar{X}\ensuremath{)} \HOLFreeVar{E}\ensuremath{)}
     \HOLKeyword{else} \HOLConst{rec} \HOLFreeVar{X} \ensuremath{(}\HOLConst{CCS_SUBST} \HOLFreeVar{fm} \HOLFreeVar{E}\ensuremath{)}\hfill{[CCS_SUBST_rec]}
\end{alltt}
Variable substitution only occurs on free
variables. In the case of $\recu X E$: if $X$ (a bound variable) is in
the domain of \HOLinline{\HOLFreeVar{fm}}, \HOLinline{\HOLConst{CCS_SUBST}} must continue the
substitution on $E$ using a reduced map~--- without $X$~--- so that all
occurences of $X$ in $E$ are correctly bypassed.
Since \HOLinline{\HOLConst{CCS_SUBST}} only runs
once on each subterm, the possible free variables in  \HOLinline{\HOLFreeVar{Ps}} are never
 substituted. 
% Thus the process is indeed
% order-independent. 
Below we present some key lemmas about
\HOLinline{\HOLConst{CCS_SUBST}}, omitting the proofs:\footnote{Hereafter,
  some new logical constants from HOL's \texttt{pred_set}
and \texttt{list} theories are used (c.f.~\cite{holdesc} for more
details.): \HOLinline{\HOLConst{DISJOINT}} denotes set disjointness;
``\HOLinline{\HOLConst{set}\\\;\HOLFreeVar{Xs}}'' is the set converted from the list \HOLinline{\HOLFreeVar{Xs}};
\HOLinline{\HOLConst{ALL_DISTINCT}} says that
the elements of a list are all distinct; \HOLinline{\HOLConst{MAP}} is
the mapping function from one list to another; \HOLinline{\HOLConst{EVERY}} means
each element of a list satisfies a predicate; \HOLinline{\HOLConst{ZIP}} creates a
new list from two lists of the same length, and each element of the new list is a pair of
elements from  the given lists.}

>>>>

Proposition~\ref{prop:transFV} \hl{is essential in the proofs of the
  multivariable versions of all unique-solution
  theorems.
In comparison with} the above proof, the proof of the analogous
result for bound variables is  easier:
\begin{proposition}
\label{prop:transBV}
if $E \arr{\mu} E'$ then $\bv{E'} \subseteq \bv{E}$, or formally:
\begin{alltt}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{E} \HOLBoundVar{u} \ensuremath{\HOLBoundVar{E}\sp{\prime}}. \HOLBoundVar{E} \HOLTokenTransBegin\HOLBoundVar{u}\HOLTokenTransEnd \ensuremath{\HOLBoundVar{E}\sp{\prime}} \HOLSymConst{\HOLTokenImp{}} \HOLConst{BV} \ensuremath{\HOLBoundVar{E}\sp{\prime}} \HOLSymConst{\HOLTokenSubset{}} \HOLConst{BV} \HOLBoundVar{E}\hfill{[TRANS_BV]}
\end{alltt}
\end{proposition}
\begin{proof}
Following the proof of Proposition~\ref{prop:transFV}, we do action
induction on $E \arr{\mu} E'$ and simplify all subgoals by set-theoretic
facts. The only subgoal left is:
\begin{alltt}
        \HOLinline{\HOLConst{BV} \ensuremath{\HOLFreeVar{E}\sp{\prime}} \HOLSymConst{\HOLTokenSubset{}} \HOLFreeVar{X} \HOLConst{INSERT} \HOLConst{BV} \HOLFreeVar{E}}
   ------------------------------------
    0.  \HOLinline{\HOLConst{BV} \ensuremath{\HOLFreeVar{E}\sp{\prime}} \HOLSymConst{\HOLTokenSubset{}} \HOLConst{BV} \ensuremath{(}\ensuremath{[}\HOLConst{rec} \HOLFreeVar{X} \HOLFreeVar{E}\ensuremath{/}\HOLFreeVar{X}\ensuremath{]} \HOLFreeVar{E}\ensuremath{)}}
\end{alltt}
Note the difference with the proof of Proposition~\ref{prop:transFV}: now
we have \texttt{INSERT} in place of \texttt{DELETE}. Then, we directly
use the assumption and the transitivity of $\subseteq$ and try to
prove the following goal instead:
\begin{alltt}
        \HOLinline{\HOLConst{BV} \ensuremath{(}\ensuremath{[}\HOLConst{rec} \HOLFreeVar{X} \HOLFreeVar{E}\ensuremath{/}\HOLFreeVar{X}\ensuremath{]} \HOLFreeVar{E}\ensuremath{)} \HOLSymConst{\HOLTokenSubset{}} \HOLFreeVar{X} \HOLConst{INSERT} \HOLConst{BV} \HOLFreeVar{E}}
   ------------------------------------
    0.  \HOLinline{\HOLConst{BV} \ensuremath{\HOLFreeVar{E}\sp{\prime}} \HOLSymConst{\HOLTokenSubset{}} \HOLConst{BV} \ensuremath{(}\ensuremath{[}\HOLConst{rec} \HOLFreeVar{X} \HOLFreeVar{E}\ensuremath{/}\HOLFreeVar{X}\ensuremath{]} \HOLFreeVar{E}\ensuremath{)}}
\end{alltt}
This is an easy goal, due to the following lemma which is
 proved by \hl{doing} induction on $E$:
\begin{alltt}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{X} \HOLBoundVar{E} \ensuremath{\HOLBoundVar{E}\sp{\prime}}. \HOLConst{BV} \ensuremath{(}\ensuremath{[}\ensuremath{\HOLBoundVar{E}\sp{\prime}}\ensuremath{/}\HOLBoundVar{X}\ensuremath{]} \HOLBoundVar{E}\ensuremath{)} \HOLSymConst{\HOLTokenSubset{}} \HOLConst{BV} \HOLBoundVar{E} \HOLSymConst{\HOLTokenUnion{}} \HOLConst{BV} \ensuremath{\HOLBoundVar{E}\sp{\prime}}\hfill{[BV_SUBSET]}
\end{alltt}
Taking $E ' = \recu X E$ in the above lemma, we have
    \HOLinline{\HOLConst{BV}\\\;\ensuremath{(}\ensuremath{[}\HOLConst{rec}\\\;\HOLFreeVar{X}\\\;\HOLFreeVar{E}\ensuremath{/}\HOLFreeVar{X}\ensuremath{]}\\\;\HOLFreeVar{E}\ensuremath{)}\\\;\HOLSymConst{\HOLTokenSubset{}}\\\;\HOLConst{BV}\\\;\HOLFreeVar{E}\\\;\HOLSymConst{\HOLTokenUnion{}}\\\;\HOLConst{BV}\\\;\ensuremath{(}\HOLConst{rec}\\\;\HOLFreeVar{X}\\\;\HOLFreeVar{E}\ensuremath{)}},
    the right side equals to \HOLinline{\HOLConst{BV}\\\;\HOLFreeVar{E}\\\;\HOLSymConst{\HOLTokenUnion{}}\\\;\ensuremath{(}\HOLFreeVar{X}\\\;\HOLConst{INSERT}\\\;\HOLConst{BV}\\\;\HOLFreeVar{E}\ensuremath{)}} $=$
    \HOLinline{\HOLFreeVar{X}\\\;\HOLConst{INSERT}\\\;\HOLConst{BV}\\\;\HOLFreeVar{E}}. This
completes
the    proof.
\end{proof}

@@

   \HOLConst{CCS_SUBST} \HOLFreeVar{fm} \ensuremath{(}\HOLConst{rec} \HOLFreeVar{X} \HOLFreeVar{E}\ensuremath{)} \HOLTokenDefEquality{}
     \HOLKeyword{if} \HOLFreeVar{X} \HOLSymConst{\HOLTokenIn{}} \HOLConst{FDOM} \HOLFreeVar{fm} \HOLKeyword{then} \HOLConst{rec} \HOLFreeVar{X} \ensuremath{(}\HOLConst{CCS_SUBST} \ensuremath{(}\HOLFreeVar{fm} \HOLSymConst{\ensuremath{\setminus}} \HOLFreeVar{X}\ensuremath{)} \HOLFreeVar{E}\ensuremath{)}
     \HOLKeyword{else} \HOLConst{rec} \HOLFreeVar{X} \ensuremath{(}\HOLConst{CCS_SUBST} \HOLFreeVar{fm} \HOLFreeVar{E}\ensuremath{)}\hfill{[CCS_SUBST_rec]}
\end{alltt}
Variable substitution only occurs on free
variables. In the case of $\recu X E$: if $X$ (a bound variable) is in
the domain of \HOLinline{\HOLFreeVar{fm}}, \HOLinline{\HOLConst{CCS_SUBST}} must continue the
substitution on $E$ using a reduced map~--- without $X$~--- so that all
occurences of $X$ in $E$ are correctly bypassed.
Since \HOLinline{\HOLConst{CCS_SUBST}} only runs
once on each subterm, the possible free variables in  \HOLinline{\HOLFreeVar{Ps}} are never
 substituted. 
% Thus the process is indeed
% order-independent. 
Below we present some key lemmas about
\HOLinline{\HOLConst{CCS_SUBST}}, omitting the proofs:\footnote{Hereafter,
  some new logical constants from HOL's \texttt{pred_set}
and \texttt{list} theories are used (see~\cite{holdesc} for more
details.): \HOLinline{\HOLConst{DISJOINT}} denotes set disjointness;
``\HOLinline{\HOLConst{set}\\\;\HOLFreeVar{Xs}}'' is the set converted from the list \HOLinline{\HOLFreeVar{Xs}};
\HOLinline{\HOLConst{ALL_DISTINCT}} says that
the elements of a list are all distinct; \HOLinline{\HOLConst{MAP}} is
the mapping function from one list to another; \HOLinline{\HOLConst{EVERY}} means
each element of a list satisfies a predicate; \HOLinline{\HOLConst{ZIP}} creates a
new list from two lists of the same length, and each element of the new list is a pair of
elements from  the given lists.}
